IFA Comparison of Noncognitive Skills \\ BCS and MCS

### 2018-03-28 codes/R/cohorts_nc.Rnw
graphs for CF decomposition

<<MAIN,cache=FALSE,echo=FALSE,results='hide',message=FALSE,warning=FALSE>>=
rm(list=ls())         # Clear all objects from memory 
  opts_chunk$set(concordance=FALSE,
                 echo=FALSE,
                 tidy=FALSE,
                 autodep=TRUE,
                 cache=T, # REPLACE LATER
                 message=FALSE, 
                 warning=FALSE,
                 size='scriptsize',
                 results='hide',
                 comment=NA, 
                 fig.width=4,
                 fig.height=4, 
                 out.width='\\textwidth',
                 fig.show='hold',
                 tikzDefaultEngine='pdftex',
                 tikzDocumentDeclaration = "\\documentclass[10pt]{article}",
                 tikzMetricPackages = c("\\usepackage[utf8]{inputenc}","\\usepackage[T1]{fontenc}", "\\usetikzlibrary{calc}", "\\usepackage{amssymb}")
                 )

@@

<<READ, cache=FALSE >>=
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/1_load.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/2_efa.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/3_mifa_invariance.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/3_mifa_final.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/4_scores.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/4_scores_compare.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/5_graphs.R")

@@

<<PREAMBLE, cache=F>>=
@

@@

<<BCSEFA>>=
@

<<BCSEFARES, dependson='BCSEFA', echo=T, results='markup'>>=
@

@@

<<MCSEFA>>=
@

<<MCSEFARES, dependson='BCSEFA', echo=T, results='markup'>>=
@

@@

\clearpage
\begin{landscape}
\begin{table}[ht]
\centering
\footnotesize
\caption{Item Commonalities}\label{tab:comm}
<<MEANTABLE_RES, dependson="LOAD_DATA", results='asis'>>=
require(xtable)

@@

A substantially identical approach is proposed for categorical measures by \cite{Millsap2004}. However, new identification results in \cite{Wu2016} indicate that, in the case of categorical measures, invariance properties cannot be examined by simply restricting one set of parameters at a time. This is because the identification conditions used in the configural baseline model, while being minimally restrictive on their own, become restrictive once certain additional restrictions are imposed. In light of this, they propose sets of restrictions that identify structures of different invariance levels. They find that some restrictions cannot be tested alone against the configural model, because they give rise to statistically equivalent models. This is true of loading invariance, and also of threshold invariance in the case when the number of categories of each ordinal item is 3 or less. Furthermore, they suggest that comparison of both latent means and variances requires invariance in loadings, thresholds, and intercepts.

@@

I estimate each of the models listed above by Weighted Least Squares.\footnote{All estimates are computed using the \texttt{lavaan} package in \textbf{\textsf{R}} \citep{Rosseel2012}. Parameters of the model are estimated by mean- and variance adjusted weighted least squares (WLSMV), due to \cite{Mplus2017}. Point estimates for model parameters are obtained starting from the polychoric correlations, using diagonally weighted least squares (DWLS), and the full weight matrix is used to compute robust standard errors and test statistics. Robust WLS has proved in simulation studies to be moderately robust to small violations of the normality assumption in the latent underlying measures \citep{Flora2011}, and generally outperforms maximum likelihood in large samples \citep{Beauducel2009,Li2016}.} The fit of the increasingly restrictive models presented above is compared in \autoref{tab:fit}. For all models, a chi-squared difference would point to a lack of measurement invariance. However, this test is know to overestimate invariance in large samples. Alternative fit indices should thus be used when comparing nested models. \cite{Cheung2002} suggest relying on Comparative Fit Index (CFI, \citealp{Bentler1990}), McDonald Non-Centrality Index (MFI, \citealp{McDonald1989}), and Gamma Hat index \citep{Steiger1989} for the case of ordered measures. Commonly accepted thresholds for rejecting the restricted model are $\Delta CFI < - 0.01$, $\Delta MFI < - 0.02$, and $\Delta Gamma hat < - 0.001$.

The overall fit of the configural model is satisfactory according to widely accepted rules of thumb \citep{Maccallum1996}.\footnote{{\color{red} BETTER CITATIONS FOR OVERALL FIT}.} As discussed above, when items have less than four categories, the model with restricted thresholds is statistically equivalent to the configural model. Restricting thresholds, loadings, and intercepts at the same time seems to provide a satisfying fit compared to the configural model.\footnote{\cite{Meade2008}, using the results from a simulation study, suggests stricter thresholds that should apply in a variety of conditions. For CFI, a single cutoff values of $.002$ is proposed, while cutoffs for MFI depend on the problem's characteristics; in our case (2 factors, 11 items), they suggest $.0066$. \cite{Sass2014} however cast some doubts of the generalisability of these cutoffs to WLSMV estimators.}

@@

\textbf{REVIEW} We can derive scores for $\theta^{EXT}$ and $\theta^{INT}$ directly, and plot them against socioeconomic measures observed in the data. Our main socioeconomic indicator is a measure of weekly total family net income in the prices of the interview year.\footnote{This measure is harmonised across cohorts and derived as part of CLOSER, which is funded by the ESRC.} Using the estimated parameters for the model with invariant thresholds and loadings (but non-invariant intercepts), I compute the scores for each individual in the sample (using the Empirical Bayes modal approach). In \autoref{fig:boxscores}, we can see that the gradient in noncognitive skills along family income quintiles is more pronounced in the MCS cohort. 
%In particular, a male five-year old at the top quintile of the income distribution in 1975 has a better score by \Sexpr{regcoeffs[1,3]} in the externalising dimension and \Sexpr{regcoeffs[1,4]} in the internalising dimension, compared to a child at the bottom quintile. In 2005, the differences amount to \Sexpr{regcoeffs[3,3]} and \Sexpr{regcoeffs[3,4]}, respectively.

@@

\clearpage

List of estimated models:
\bnm
\item no gender split, no age adjustment (2 groups, BCS MCS)
\item no gender split, age adjustment (2 groups, BCS MCS)
\item males only, no age adjustment (2 groups, BCS MCS)
\item females only, no age adjustment (2 groups, BCS MCS)
\item males only, age adjustment (2 groups, BCS MCS)
\item females only, age adjustment (2 groups, BCS MCS)
\item separate gender groups, no age adjustment (4 groups, BCS.M BCS.F MCS.M MCS.F)
\item separate gender groups, age adjustment (4 groups, BCS.M BCS.F MCS.M MCS.F)
\item separate gender groups, overlapping ages, no age adjustment (4 groups, BCS.M BCS.F MCS.M MCS.F)
\item separate gender groups, age adjustment constrained to be the same across ages (4 groups, BCS.M BCS.F MCS.M MCS.F)
\item MIMIC model with age as regressor (4 groups, BCS.M BCS.F MCS.M MCS.F)
\enm

Notes: 
\bi
\item Only the models with age adjustment seem to have invariant intercepts (see Appendix \autoref{sec:app_fit}). 
\item But the latent means of these models don't make sense with the "raw" scores we have (see Appendix \autoref{sec:app_meancov})
\item Also, a model on the subset of children that have very similar ages delivers results that are very similar to the age-unadjusted models, and different from the age-adjusted ones. So the models with age adjustment seem to be doing something else than adjusting for age, and we don't understand what this something else is...
\item The model with free intercepts (only restricting thresholds and loadings) delivers differences in scores by group, even if the means of the latent factors are constrained to be equal (see below)
\item The models with free and constrained intercepts deliver very similar scores (.98 correlation), and they also give sensible and comparable gradients with income
\ei

<<FA_INVARIANCE, dependson='LOAD_DATA'>>=
# estimate models of different levels of invariance
@

<<FA_FIT, dependson='FA_INVARIANCE'>>=
# assemble fit indices into matrices
@

<<FA_FINAL, dependson='FA_INVARIANCE'>>=
# select final model set
@

<<FA_FINAL_PARS, dependson='FA_FINAL'>>=
# get parameters of final model
@

<<CLEANUP, dependson='FA_FINAL'>>=
@

<<FA_SCORES, dependson='FA_FINAL'>>=
# produce scores for different models
@

<<FA_SCORES_COMPARE, dependson='FA_SCORES'>>=
# assemble and compare scores across different models
@

<<FA_SCORES_SELECT, dependson='FA_FINAL'>>=
# select model to use for inequality analysis
@

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<FA_SCORES_COMPARE, dependson='FA_SCORES'>>=
# assemble and compare scores across different models

When adjusting for age, mean scores do not agree with raw scores
<<FA_SCORES_MEANS, dependson='FA_SCORES', echo=T, results='markup'>>=
# means
means.ext
means.int
@

Scores that do not adjust for age are very highly correlated, regardless of whether they come from a model with free or restricted intercepts
<<FA_SCORES_CORR, dependson='FA_SCORES', echo=T, results='markup'>>=
cor(compare.scores[,c(grep("EXT", names(compare.scores), value=T))])
cor(compare.scores[,c(grep("INT", names(compare.scores), value=T))])
@

The gradient of scores by income quintile are very similar
<<FA_SCORES_INEQ, dependson='FA_SCORES', echo=T, results='markup'>>=

# free intercepts
lm(EXTtl ~ factor(incq), data=compare.scores)$coefficients
lm(INTtl ~ factor(incq), data=compare.scores)$coefficients

# constrained intercepts
lm(EXTtli ~ factor(incq), data=compare.scores)$coefficients
lm(INTtli ~ factor(incq), data=compare.scores)$coefficients

# constrained intercepts, common ages
lm(EXTtlir ~ factor(incq), data=compare.scores)$coefficients
lm(INTtlir ~ factor(incq), data=compare.scores)$coefficients

<<FA_REGS, dependson='FA_SCORES'>>=

# combinations of regressions
regcoeffs <- matrix( 
  c(rep("BCS", 2), rep("MCS", 2), rep(c("M","F"), 2), rep(NA, 8)), # the data elements 
  nrow=4,              # number of rows 
  ncol=4,              # number of columns 
  byrow = F)        # fill matrix by columns 
colnames(regcoeffs) <- c("cohort", "sex", "EXTd", "INTd")

# fill in differences
regs.ext <- list()
regs.int <- list()
for (r in 1:nrow(regcoeffs)) {
  regs.ext[[r]] <- lm(EXT ~ as.factor(incq) + age, data=subset(scores2plot, cohort==regcoeffs[r,1] & sex==regcoeffs[r,2]))
  regcoeffs[r,3] <- round(regs.ext[[r]]$coefficients["as.factor(incq)5"],2)
  regs.int[[r]] <- lm(INT ~ as.factor(incq) + age, data=subset(scores2plot, cohort==regcoeffs[r,1] & sex==regcoeffs[r,2]))
  regcoeffs[r,4] <- round(regs.int[[r]]$coefficients["as.factor(incq)5"],2)
}

% \begin{table}[ht]
% \caption{Estimated means of scores -- non-age adjusted 4-group model} \label{tab:scoremeans}
% \centering
% <<FA_COMP1, dependson='FA_SCORES', results='asis'>>=
% # compare mean factor scores by gender and cohort
% 
% prn <- function(obj) sprintf("%.3f", round(obj,3))
% toaggr <- c("EXT", "INT", "EXTr", "INTr", "EXT.RAW", "INT.RAW", "EXT.RAWr", "INT.RAWr")
% sctable <- aggregate(items.scored[[7]][, toaggr], list(items.scored[[7]]$cohortsex), 
%                    FUN = function (x) prn(mean(x)))
% 
% sctable <- cbind(group = factor(c(1,3,2,4)), sctable)
% levels(sctable$group) <- c("BCS Males", "MCS Males", "BCS Females", "MCS Females")
% sctable <- sctable[order(sctable$group),]
% sctable <- sctable[, !names(sctable) %in% "Group.1"]
% sctable <- as.matrix(sctable,3)
% 
% 
% print.xtable(xtable(sctable,
%                     align="L{0cm}L{2.2cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}",
%                     digits = 3
%                     ),
%              size="\\small",
%              include.rownames=F,
%              include.colnames=F,
%              booktabs = T,
%              add.to.row = list(pos = c(list(0,4)),
%                               command = c(
%                               # header
%                                    "\\toprule
%                                     & \\multicolumn{4}{c}{\\textbf{Scored factors}}
%                                     & \\multicolumn{4}{c}{\\textbf{Scales}} \\\\ \\cmidrule(lr){2-5} \\cmidrule(lr){6-9}
%                                     & \\multicolumn{2}{c}{Normal} & \\multicolumn{2}{c}{Age-adjusted} & \\multicolumn{2}{c}{Normal} & \\multicolumn{2}{c}{Age-Adjusted} \\\\ \\cmidrule(lr){2-3} \\cmidrule(lr){4-5} \\cmidrule(lr){6-7} \\cmidrule(lr){8-9}
%                       & EXT & INT & EXT & INT & EXT & INT & EXT & INT \\\\",
%                       "\\bottomrule"
%                                     )),
%              hline.after = NULL,  # remove default booktabs lines
%              sanitize.text.function = function(x){x},
%              sanitize.rownames.function = function(x){""}, # to not print row numbers
%              tabular.environment = "tabularx", width = "\\textwidth",
%              floating = FALSE
% )
% 
% @
% \end{table}
% 
% \begin{table}[ht]
% \caption{Estimated means of scores -- age adjusted 4-group model} \label{tab:scoremeans}
% \centering
% <<FA_COMP2, dependson='FA_SCORES', results='asis'>>=
% # compare mean factor scores by gender and cohort
% 
% prn <- function(obj) sprintf("%.3f", round(obj,3))
% toaggr <- c("EXT", "INT", "EXTr", "INTr", "EXT.RAW", "INT.RAW", "EXT.RAWr", "INT.RAWr")
% sctable <- aggregate(items.scored[[8]][, toaggr], list(items.scored[[8]]$cohortsex), 
%                    FUN = function (x) prn(mean(x)))
% 
% sctable <- cbind(group = factor(c(1,3,2,4)), sctable)
% levels(sctable$group) <- c("BCS Males", "MCS Males", "BCS Females", "MCS Females")
% sctable <- sctable[order(sctable$group),]
% sctable <- sctable[, !names(sctable) %in% "Group.1"]
% sctable <- as.matrix(sctable,3)
% 
% 
% print.xtable(xtable(sctable,
%                     align="L{0cm}L{2.2cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}",
%                     digits = 3
%                     ),
%              size="\\small",
%              include.rownames=F,
%              include.colnames=F,
%              booktabs = T,
%              add.to.row = list(pos = c(list(0,4)),
%                               command = c(
%                               # header
%                                    "\\toprule
%                                     & \\multicolumn{4}{c}{\\textbf{Scored factors}}
%                                     & \\multicolumn{4}{c}{\\textbf{Scales}} \\\\ \\cmidrule(lr){2-5} \\cmidrule(lr){6-9}
%                                     & \\multicolumn{2}{c}{Normal} & \\multicolumn{2}{c}{Age-adjusted} & \\multicolumn{2}{c}{Normal} & \\multicolumn{2}{c}{Age-Adjusted} \\\\ \\cmidrule(lr){2-3} \\cmidrule(lr){4-5} \\cmidrule(lr){6-7} \\cmidrule(lr){8-9}
%                       & EXT & INT & EXT & INT & EXT & INT & EXT & INT \\\\",
%                       "\\bottomrule"
%                                     )),
%              hline.after = NULL,  # remove default booktabs lines
%              sanitize.text.function = function(x){x},
%              sanitize.rownames.function = function(x){""}, # to not print row numbers
%              tabular.environment = "tabularx", width = "\\textwidth",
%              floating = FALSE
% )
% 
% @
% \end{table}
% 
% \begin{table}[ht]
% \caption{Estimated means of scores -- 4-group model, 59-62 months} \label{tab:scoremeans}
% \centering
% <<FA_COMP3, dependson='FA_SCORES', results='asis'>>=
% # compare mean factor scores by gender and cohort
% 
% prn <- function(obj) sprintf("%.3f", round(obj,3))
% toaggr <- c("EXT", "INT", "EXTr", "INTr", "EXT.RAW", "INT.RAW", "EXT.RAWr", "INT.RAWr")
% sctable <- aggregate(items.scored[[9]][, toaggr], list(items.scored[[9]]$cohortsex), 
%                    FUN = function (x) prn(mean(x)))
% 
% sctable <- cbind(group = factor(c(1,3,2,4)), sctable)
% levels(sctable$group) <- c("BCS Males", "MCS Males", "BCS Females", "MCS Females")
% sctable <- sctable[order(sctable$group),]
% sctable <- sctable[, !names(sctable) %in% "Group.1"]
% sctable <- as.matrix(sctable,3)
% 
% 
% print.xtable(xtable(sctable,
%                     align="L{0cm}L{2.2cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}",
%                     digits = 3
%                     ),
%              size="\\small",
%              include.rownames=F,
%              include.colnames=F,
%              booktabs = T,
%              add.to.row = list(pos = c(list(0,4)),
%                               command = c(
%                               # header
%                                    "\\toprule
%                                     & \\multicolumn{4}{c}{\\textbf{Scored factors}}
%                                     & \\multicolumn{4}{c}{\\textbf{Scales}} \\\\ \\cmidrule(lr){2-5} \\cmidrule(lr){6-9}
%                                     & \\multicolumn{2}{c}{Normal} & \\multicolumn{2}{c}{Age-adjusted} & \\multicolumn{2}{c}{Normal} & \\multicolumn{2}{c}{Age-Adjusted} \\\\ \\cmidrule(lr){2-3} \\cmidrule(lr){4-5} \\cmidrule(lr){6-7} \\cmidrule(lr){8-9}
%                       & EXT & INT & EXT & INT & EXT & INT & EXT & INT \\\\",
%                       "\\bottomrule"
%                                     )),
%              hline.after = NULL,  # remove default booktabs lines
%              sanitize.text.function = function(x){x},

>>>>

<<MAIN,cache=FALSE,echo=FALSE,results='hide',message=FALSE,warning=FALSE>>=
rm(list=ls())         # Clear all objects from memory 
  opts_chunk$set(concordance=FALSE,
                 echo=FALSE,
                 tidy=FALSE,
                 autodep=TRUE,
                 cache=TRUE,
                 message=FALSE, 
                 warning=FALSE,
                 size='scriptsize',
                 results='hide',
                 comment=NA, 
                 fig.width=4,
                 fig.height=4, 
                 out.width='\\textwidth',
                 fig.show='hold',
                 tikzDefaultEngine='pdftex',
                 tikzDocumentDeclaration = "\\documentclass[10pt]{article}",
                 tikzMetricPackages = c("\\usepackage[utf8]{inputenc}","\\usepackage[T1]{fontenc}", "\\usetikzlibrary{calc}", "\\usepackage{amssymb}")
                 )

@@

<<READ, cache=FALSE >>=
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/1_load.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/2_efa.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/3_inv_main.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/3_inv_aux.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/3_inv_fit.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/4_scores.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/4_scores_aux.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/5_graphs.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/6_ineq.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/7_counterf.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/7_counterf_graphs.R")

@@

<<PREAMBLE, cache=FALSE >>=
@

@@

<<BCSEFA>>=
@
% 
% <<BCSEFARES, dependson='BCSEFA', echo=T, results='markup'>>=
% @

@@

<<MCSEFA>>=
@
% 
% # <<MCSEFARES, dependson='BCSEFA', echo=T, results='markup'>>=
% # @
%

@@

\clearpage
\begin{landscape}
\begin{table}[ht]
\centering
\footnotesize
\caption{Item Commonalities}\label{tab:comm}
<<MEANTABLE_RES, dependson="MEANTABLE", results='asis'>>=
require(xtable)

@@

A substantially similar approach is proposed for categorical measures by \cite{Millsap2004}. However, new identification results in \cite{Wu2016} indicate that, in the case of categorical measures, invariance properties cannot be examined by simply restricting one set of parameters at a time. This is because the identification conditions used in the configural baseline model, while being minimally restrictive on their own, become restrictive once certain additional restrictions are imposed. In light of this, they propose sets of restrictions that identify structures of different invariance levels. They find that some restrictions cannot be tested alone against the configural model, because they give rise to statistically equivalent models. This is true of loading invariance, and also of threshold invariance in the case when the number of categories of each ordinal item is 3 or less. Furthermore, they suggest that comparison of both latent means and variances requires invariance in loadings, thresholds, and intercepts.

@@

I estimate each of the models listed above by Weighted Least Squares.\footnote{All estimates are computed using the \texttt{lavaan} package in \textbf{\textsf{R}} \citep{Rosseel2012}. Parameters of the model are estimated by mean- and variance adjusted weighted least squares (WLSMV) (see \citealp{Mplus2017}). Point estimates for model parameters are obtained starting from the polychoric correlations, using diagonally weighted least squares (DWLS), and the full weight matrix is used to compute robust standard errors and test statistics. Robust WLS has proved in simulation studies to be moderately robust to small violations of the normality assumption in the latent underlying measures \citep{Flora2011}, and generally outperforms maximum likelihood in large samples \citep{Beauducel2009,Li2016}.} The fit of the increasingly restrictive models presented above is compared in \autoref{tab:fit}. For all models, a chi-squared difference would point to a lack of measurement invariance. However, this test is know to overestimate invariance in large samples. Alternative fit indices should thus be used when comparing nested models. \cite{Cheung2002} suggest relying on Comparative Fit Index (CFI, \citealp{Bentler1990}), McDonald Non-Centrality Index (MFI, \citealp{McDonald1989}), and Gamma Hat index \citep{Steiger1989} for the case of ordered measures. Commonly accepted thresholds for rejecting the restricted model are $\Delta CFI < - 0.01$, $\Delta MFI < - 0.02$, and $\Delta Gamma hat < - 0.001$.

The overall fit of the configural model is satisfactory according to widely accepted rules of thumb \citep{Maccallum1996}.\footnote{{\color{red} BETTER CITATIONS FOR OVERALL FIT}.} As discussed above, when items have less than four categories, the model with restricted thresholds is statistically equivalent to the configural model.\footnote{\cite{Meade2008}, using the results from a simulation study, suggests stricter thresholds that should apply in a variety of conditions. For CFI, a single cutoff values of $.002$ is proposed, while cutoffs for MFI depend on the problem's characteristics; in our case (2 factors, 11 items), they suggest $.0066$. \cite{Sass2014} however cast some doubts of the generalisability of these cutoffs to WLSMV estimators.}

{\color{red} COMMENT FIT RESULTS}

<<FA_INV_MAIN, dependson="LOAD_DATA">>=
# main models only
@

% % <<FA_INV_AUX, dependson='LOAD_DATA'>>=
% % # robustness models
% % @

<<FA_FIT, dependson='FA_INV_MAIN'>>=
# assemble fit indices into matrices
@

<<FA_SCORES_MAIN, dependson='FA_INV_MAIN'>>=
# produce scores for main models
@

% % <<FA_SCORES_AUX, dependson='FA_INV_AUX'>>=
% % # produce scores for robustness models
% % @

\begin{table}[ht]
\centering
\small
\caption{Measurement invariance fit comparison}\label{tab:fit}
<<FA_FIT_RES, dependson="FA_FIT", results='asis'>>=
require(xtable)

table <- rbind(afitab[[1]],afitab[[2]])

print(xtable(table,
                    align="L{0cm}L{.20\\textwidth}C{.03\\textwidth}C{.07\\textwidth}C{.07\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}",
                    digits = 4
                    ),
             size="\\small",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,3)),
                              command = c(
                              # header
            "Model & Num. params & $\\chi^2$ & RMSEA & MFI & CFI & G-hat & \\textDelta MFI & \\textDelta CFI & \\textDelta G-hat \\\\ \\midrule
                       \\textbf{All ages} \\\\",
                      "\\newline \\textbf{59-61 months sample} \\\\ \\midrule "
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)

\clearpage
\begin{figure}
<<FACDENS, dependson="FA_SCORES_MAIN", dev='tikz', fig.height=10, fig.width=10, out.width='1\\linewidth'>>=
@
\caption{Density of Scored Factors}
{\scriptsize \emph{Notes}: The figure shows estimated kernel densities of the scored factors separately by cohort.  \par}
\end{figure}

@@

\textbf{REVIEW} We can derive scores for $\theta^{EXT}$ and $\theta^{INT}$ directly, and plot them against socioeconomic measures observed in the data. Our main socioeconomic indicator is a measure of weekly total family net income in the prices of the interview year.\footnote{This measure is harmonised across cohorts and derived as part of CLOSER, which is funded by the ESRC.} Using the estimated parameters for the model with invariant thresholds and loadings (but non-invariant intercepts), I compute the scores for each individual in the sample (using the Empirical Bayes modal approach). In \autoref{fig:meanci}, we can see that the gradient in noncognitive skills along family income quintiles is more pronounced in the MCS cohort.
%In particular, a male five-year old at the top quintile of the income distribution in 1975 has a better score by \Sexpr{regcoeffs[1,3]} in the externalising dimension and \Sexpr{regcoeffs[1,4]} in the internalising dimension, compared to a child at the bottom quintile. In 2005, the differences amount to \Sexpr{regcoeffs[3,3]} and \Sexpr{regcoeffs[3,4]}, respectively.

\begin{figure}
<<FACINEQ, dependson="FA_SCORES_MAIN", dev='tikz', fig.height=12, fig.width=10, out.width='1\\linewidth'>>=
@
\caption{Mean scores across income quintiles}\label{fig:meanci}
{\scriptsize \emph{Notes}: The figure plots the mean and its 95\% confidence interval across the quintiles in the income distribution. \par}
\end{figure}

\begin{figure}
<<FACINEQ_SC, dependson="FA_SCORES_MAIN", dev='tikz', fig.height=12, fig.width=10, out.width='1\\linewidth'>>=
@
\caption{Mean scores across social class}\label{fig:meanci_sc}
{\scriptsize \emph{Notes}: The figure plots the mean and its 95\% confidence interval across the quintiles in the income distribution. \par}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage
% THIS BREAKS THE CACHING FOR SOME REASON
% <<IQRDEC_FUN>>=
% # load functions
% @
% 
% <<IQRDEC, dependson="FA_SCORES_MAIN">>=
% # decompose
% @
% 
% % Compare inequality statistics of original vs residualised scores
% % <<IQRDEC_RES, dependson='IQRDEC', echo=T, results='markup'>>=
% % iqs.orig
% % iqs.resid
% % @

\clearpage
\subsection{Decomposition of inequality}

We have populations $k \in \mathcal{K}$. For each population, we observe covariates $X_k$. For some $j \subseteq \mathcal{K}$, we observe outcome $Y_j$. We can identify the covariate distributions $F_{X_k}$ and the conditional distributions $F_{Y_j \mid X_j}$ and quantile functions $Q_{Y_j \mid X_j}$. Define the support of $X_k$ as $\mathcal{X}_k$ and the region of interest for the outcome as $\mathcal{Y}_j$.

The interest is in the counterfactual distribution of the outcome:
$$ F_{Y(j \mid k)}(y) = \int_{\mathcal{X}_k} F_{Y_j \mid X_j}(y \mid x) d F_{X_k}(x), \qquad y \in \mathcal{Y}_j $$
and the corresponding quantiles:
$$ Q_{Y(j \mid k)}(\tau) = F^\leftarrow_{Y(j \mid k)}(\tau) $$

Given this, we can define counterfactual effects (CE):
\bi
\item Type 1 (Structure effect): CE of changing the conditional distribution: $Q_{Y(j \mid k)}(\tau) - Q_{Y(l \mid k)}(\tau)$
\item Type 2 (Composition effect): CE of changing the covariate distribution: $Q_{Y(j \mid k)}(\tau) - Q_{Y(j \mid m)}(\tau)$
\item Type 3: CE of changing both conditional and covariate distribution: $Q_{Y(j \mid k)}(\tau) - Q_{Y(l \mid m)}(\tau)$
\ei

In our application, we have $k=BCS,MCS$. We can thus decompose the observed differences in the distribution of non-cognitive skills into structure and composition effects. For example, the difference in the distribution of externalising skills can be written as:
$$F_{(\theta^{EXT}_{MCS} \mid X_{MCS})} - F_{(\theta^{EXT}_{BCS} \mid X_{BCS})} =
  \left[ F_{(\theta^{EXT}_{MCS} \mid X_{MCS})} - F_{(\theta^{EXT}_{BCS} \mid X_{MCS})} \right] +
  \left[ F_{(\theta^{EXT}_{BCS} \mid X_{MCS})} - F_{(\theta^{EXT}_{BCS} \mid X_{BCS})} \right]$$

<<COUNTERF, dependson="FA_SCORES_MAIN">>=
# perform counterfactuals
@

<<COUNTERF_PREP_PLOTS, dependson='COUNTERF'>>=
# prepare counterfactual plots
@

\begin{landscape}
\begin{figure}
<<COUNTERF_EXT, dependson='COUNTERF_PREP_PLOTS', dev='tikz', fig.height=10, fig.width=16, out.width='1\\linewidth'>>=
cfp.ext
@
\caption{Counterfactual decomposition -- externalising skills}\label{fig:counterf}
{\scriptsize \emph{Notes}: \par}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
<<COUNTERF_INT, dependson='COUNTERF_PREP_PLOTS', dev='tikz', fig.height=10, fig.width=16, out.width='1\\linewidth'>>=
cfp.int
@
\caption{Counterfactual decomposition -- internalising skills}\label{fig:counterf}
{\scriptsize \emph{Notes}: \par}
\end{figure}
\end{landscape}

@@

\clearpage
\bibliographystyle{chicago}
\bibliography{/Users/giacomomason/Documents/BibTeX/Papers-cohorts_nc}
%\bibliography{/Users/giacomomason/Dropbox/Zotero/cohorts.bib}

\clearpage
\begin{appendices}
\section{Estimated model parameters}

<<FUN_GETPARS>>=
# load function that gets parameters
@

\begin{table}[ht]
\caption{Estimated loadings and thresholds} \label{tab:allpars}
\resizebox{\textwidth}{!}{
<<RES_FA_PAR, dependson="FA_FIT", results='asis'>>=
source("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/getpars.R")

table <- getmeaspars(finalmod, groups = 4, mode = "tl")

print.xtable(xtable(table,
                    # align = "ll|cccc|cccc"
              align="C{0cm}C{1.2cm}C{1cm}C{1.3cm}C{1.1cm}C{1.1cm}C{1.1cm}C{1.1cm}C{1.1cm}C{1.1cm}C{1.1cm}C{1.1cm}",
              digits = 3
                    ),
             size="\\small",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0)),
                              command = c(
                              # header
                                          " &
                                            & \\textbf{Loadings}
                                            & \\multicolumn{2}{c}{\\textbf{Thresholds}}
                                            & \\multicolumn{3}{c}{\\textbf{Intercepts}}
                                            & \\multicolumn{3}{c}{\\textbf{Variances}}
                                            \\\\ \\cmidrule(lr){3-3} \\cmidrule(lr){4-5} \\cmidrule(lr){6-8} \\cmidrule(lr){9-11}

                                            &
                                            & $\\lambda$
                                            & $\\tau_1$
                                            & $\\tau_2$
                                            & \\multicolumn{3}{c}{$\\nu$}
                                            & \\multicolumn{3}{c}{$\\text{diag}(\\Psi)$} \\\\

                                            Measure
                                            & Factor
                                            & All
                                            & All
                                            & All
                                            & BCS F
                                            & MCS M
                                            & MCS F
                                            & BCS F
                                            & MCS M
                                            & MCS F
                                            \\\\ "
                      )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabularx", width = "\\textwidth",
             floating = FALSE
)
@
}
\end{table}

\begin{table}[ht]
\caption{Estimated latent variable means and covariances} \label{tab:lvpars}
\centering
<<RES_FA_LV, dependson="FA_FIT", results='asis'>>=
require(xtable)
source("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/getpars.R")

lvtab <- getlvpars(finalmod, groups = 4)
table <- rbind(lvtab$Males,lvtab$Females)

print.xtable(xtable(table,
                    align="L{0cm}L{3cm}C{1.2cm}C{1.2cm}C{1.2cm}C{1.2cm}C{1.2cm}C{1.2cm}C{1.2cm}C{1.2cm}",
                    digits = 3
                    ),
             size="\\small",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,2,4)),
                              command = c(
                              # header
                                   "\\toprule
                                    & \\multicolumn{4}{c}{\\textbf{BCS}}
                                    & \\multicolumn{4}{c}{\\textbf{MCS}} \\\\ \\cmidrule(lr){2-5} \\cmidrule(lr){6-9}
                                    & Mean & \\multicolumn{2}{c}{Covariance} & Corr. & Mean & \\multicolumn{2}{c}{Covariance} & Corr. \\\\ \\midrule
                      \\textbf{Males} \\\\",
                      "\\textbf{Females} \\\\",
                      "\\bottomrule"
                                    )),
             hline.after = NULL,  # remove default booktabs lines
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabularx", width = "\\textwidth",
             floating = FALSE
)
